---
title: 'Prophet (time series)'
author: 'Charlie Kopp'
date: today
date-format: long
format:
  html:
    toc: true
    code-fold: true
    embed-resources: true
    fig-align: center
execute:
  echo: true
  warning: false
  message: false

---

```{r}
#| label: setup
#| include: false

# install.packages(c("prophet", "data.table"))
library(data.table)
```

# Data

```{r}
#| label: read data

dt <- fread("STATS M148/m148-project.nosync/data/dat_train1.csv")
```

***Thomas Feedback:***

- predicting proportion of success based on first action date is too complicated. just make outcome the total orders shipped on any given day.
- also need to figure out how prophet handles missing data (no orders are shipped on weekends)

## Total orders shipped per-day approach

```{r}
#| label: format data

# just want to summarize data for total orders shipped per-day

daily_data <- dt[, .(
  y = sum(event_name=="order_shipped")
), by = .(ds = as.IDate(event_timestamp))]

# first couple weeks are all ZERO! in-progress orders were excluded
# need to find a later start point - get avg length of successful journey and take pctile

# any duplicate rows?
anyDuplicated(prophet_train_data$ds)
```

## Total orders shipped per-month

```{r}
#| label: monthly orders data

prophet_train_data <- dt[
  event_name == "order_shipped", 
  .(ds = as.IDate(event_timestamp))] |>
  _[, .(y = .N), by = .(ds = round(ds, "months"))]
```

## First-action date p-success approach

```{r}
#| label: format data (old)

# drop incomplete rows

train_cutoff_date <- as.IDate(max(dt$event_timestamp))

# flatten data by user

user_dt <- dt[, .(
    first_action_date = as.IDate(min(event_timestamp)),
    last_action_date = as.IDate(max(event_timestamp)),
    has_shipped = any(event_name == "order_shipped"),
    total_actions = .N
), by = id]

# define days since last action as days_inactive

user_dt[, days_inactive := as.integer(train_cutoff_date - last_action_date)]

user_dt[, journey_outcome := fcase(
    has_shipped == TRUE, "success",
    has_shipped == FALSE & days_inactive > 60, "failure",
    default = "incomplete"
)]

table(user_dt$journey_outcome)

# drop incomplete journeys

user_dt_complete <- user_dt[journey_outcome != "incomplete", ]

# want to find distr. journey length for a successes and add upper end to cutoff date to reduce survivorship bias

successes_only <- user_dt_complete[journey_outcome == "success"]

purchase_times <- dt[event_name == "order_shipped", 
                     .(purchase_date = as.IDate(min(event_timestamp))), 
                     by = id]

# join back to find the journey length
journey_lengths <- merge(successes_only, purchase_times, by = "id")
journey_lengths[, days_to_buy := as.integer(purchase_date - first_action_date)]

hist(journey_lengths$days_to_buy, breaks = 100, freq=FALSE)

# find the 95th percentile of how long it takes people to buy
p95_days_success <- quantile(journey_lengths$days_to_buy, probs = 0.95, na.rm = TRUE)
print(paste("95% of successful users buy within", round(p95_days_success), "days"))

p80_days_success <- quantile(journey_lengths$days_to_buy, probs = 0.80, na.rm = TRUE)

# define safe cohort start date
# Safe Date = Cutoff - (95th Percentile Days + 60 Days Inactivity)
safe_cutoff <- train_cutoff_date - (round(p95_days_success) + 60)

cohort_users <- user_dt_complete[first_action_date <= safe_cutoff]

prophet_train_data <- cohort_users[, .(
    total_users = .N,
    successes = sum(journey_outcome == "success"),
    y = sum(journey_outcome == "success") / .N # Prophet requires the target to be named 'y'
), by = .(ds = first_action_date)] # Prophet requires the date to be named 'ds'
```

# Model fitting

## fitting to orders-per-day

```{r}
#| label: fit prophet model to daily data

library(prophet)

# Initialize Prophet for daily data
m <- prophet(
  yearly.seasonality = TRUE,
  weekly.seasonality = TRUE,
  daily.seasonality = FALSE
)
m <- add_country_holidays(m, country_name = "US")
m <- fit.prophet(m, prophet_train_data)

# forecast through end of 2024:

# calculate days needed to reach Dec 2024
last_ds <- max(prophet_train_data$ds)
n_days <- length(seq(last_ds, as.IDate("2024-12-01"), by = "day")) - 1

# Create future dataframe at daily frequency
future_daily <- make_future_dataframe(m, periods = max(0, n_days), freq = "day")
forecast_daily <- predict(m, future_daily)
```

## fitting to orders-per-month

```{r}
#| label: fit prophet model to monthly data

library(prophet)

# Initialize Prophet for monthly data
m <- prophet(
  yearly.seasonality = TRUE,
  weekly.seasonality = FALSE,
  daily.seasonality = FALSE
)

m <- fit.prophet(m, prophet_train_data)

# Calculate months needed to reach Dec 2024
last_ds <- max(prophet_train_data$ds)
n_months <- length(seq(last_ds, as.IDate("2024-12-01"), by = "month")) - 1

# Create future dataframe at monthly frequency
future_monthly <- make_future_dataframe(m, periods = max(0, n_months), freq = "month")
forecast_monthly <- predict(m, future_monthly)
```

## fitting to cohort method

```{r}
#| label: fit prophet model

library(prophet)

m <- prophet(yearly.seasonality = TRUE, weekly.seasonality=TRUE)
m <- add_country_holidays(m, country_name = "US")
m <- fit.prophet(m, prophet_train_data)

# calculate exactly how many days of "incomplete" cohorts we need to forecast
days_to_forecast <- as.numeric(train_cutoff_date - safe_cutoff)

# create a dataframe containing both historical dates AND future dates
future_dates <- make_future_dataframe(m, periods = days_to_forecast)

# generate the forecast
# This creates a massive dataframe containing the prediction (yhat) and all components
forecast <- predict(m, future_dates)
```

# Visualiations

```{r}
#| label: plotting

library(ggplot2)
library(dplyr)
library(patchwork)
```

### daily orders model

```{r}
#| label: daily plot

plot(m, forecast_daily) +
  labs(title = "Daily Orders Forecast through 2024", x = "Date", y = "Orders Shipped") +
  theme_minimal()

ggsave("prophet-plots/daily_orders_forecast.png", width = 8, height = 4)

# components

comp_plot_list <- prophet_plot_components(m, forecast_daily)

wrap_plots(comp_plot_list, ncol = 1, nrow=4) +
  theme_minimal()

ggsave("prophet-plots/daily_orders_components.png", width = 8, height = 5)

yoy_holiday_data <- forecast_daily %>%
  select(ds, holidays) %>%
  mutate(
    # Extract the actual year to use as our color category
    Year = as.factor(year(ds)),
    
    # Force every date to occur in 2024 (a leap year to protect Feb 29)
    # This aligns 2022, 2023, and 2024 exactly on top of each other
    dummy_date = update(ds, year = 2023)
  )

ggplot(yoy_holiday_data, aes(x = dummy_date, y = holidays, color = Year)) +
  # Baseline zero line
  geom_hline(yintercept = 0, color = "darkgray", linetype = "dashed") +
  
  # Draw the lines (alpha = 0.7 makes overlapping years easier to see through)
  geom_line(linewidth = .5, alpha = 0.5) +
  
  # # Add dots only on the actual holidays
  # geom_point(data = yoy_holiday_data %>% filter(holidays != 0), size = 1) +
  
  # Force the X-axis to label every single month using the dummy dates
  scale_x_datetime(date_breaks = "1 month", date_labels = "%b") +

  scale_color_brewer() +
  
  labs(
    title = "Year-Over-Year Holiday Effect on Orders Shipped",
    subtitle = "Baseline effect during US National Holidays",
    x = "Month",
    y = "Holiday Effect",
    color = "Year"
  ) +
  theme_minimal()
```

## monthly model

```{r}
#| label: monthly plot

library(ggplot2)
library(dplyr)

# Visualize monthly forecast
plot(m, forecast_monthly) +
  labs(title = "Monthly Orders Forecast through 2024", x = "Month", y = "Orders Shipped") +
  theme_minimal()

# Plot 1: The Main Forecast 
plot(m, forecast) +
  ggtitle("FingerHut Cohort Success Rate Forecast") +
  xlab("Cohort Start Date") + 
  ylab("Baseline Success Proportion") +
  theme_minimal()

# Plot 2: The Components Breakdown
prophet_plot_components(m, forecast)

# plot 3: zoom in on holiday year-over-year

yoy_holiday_data <- forecast %>%
  select(ds, holidays) %>%
  mutate(
    # Extract the actual year to use as our color category
    Year = as.factor(year(ds)),
    
    # Force every date to occur in 2024 (a leap year to protect Feb 29)
    # This aligns 2022, 2023, and 2024 exactly on top of each other
    dummy_date = update(ds, year = 2023)
  )

ggplot(yoy_holiday_data, aes(x = dummy_date, y = holidays, color = Year)) +
  # Baseline zero line
  geom_hline(yintercept = 0, color = "darkgray", linetype = "dashed") +
  
  # Draw the lines (alpha = 0.7 makes overlapping years easier to see through)
  geom_line(linewidth = .5, alpha = 0.7) +
  
  # Add dots only on the actual holidays
  geom_point(data = yoy_holiday_data %>% filter(holidays != 0), size = 1) +
  
  # Force the X-axis to label every single month using the dummy dates
  scale_x_datetime(date_breaks = "1 month", date_labels = "%b") +
  
  labs(
    title = "Year-Over-Year Holiday Effect on Cohort Quality",
    subtitle = "Baseline success rate multiplier during US National Holidays",
    x = "Month",
    y = "Holiday Effect",
    color = "Cohort Year"
  ) +
  theme_minimal()
```