---
title: Assignment 1
subtitle: STATS M148 FingerHut Project
author: Charlie Kopp, Isaiah C. Mireles, Joel Yoon
date: today
date-format: long
format:
  pdf:
    toc: false
    number-sections: false
    geometry:
        - top=20mm
        - left=20mm
        - right=20mm
        - heightrounded
    include-in-header:
      text: |
        \usepackage{float}
        \floatplacement{table}{H}
    execute:
      echo: false
      warning: false
      message: false
      include: false
    fig-align: center
    fig-height: 6
    fig-width: 8
---

```{r}
#| label: load packages
#| include: false
library(tidyverse)
```


# Task 0 - Fingerhut

[Generated GPT](https://chatgpt.com/g/g-69cee2d550f4819195683c49639b978c-fingerhut-gpt)

[Course Notes](https://docs.google.com/document/d/1k-IZ0ObDrfif1OiEVqaGXPZ5CbQxXZzn_GQ0rBYsBRE/edit?usp=sharing)

# Task 1 - Read data

1. How many rows does your data set have?

> 54,960,961

2. How many unqiue Ids are there?

> 1,430,445

3. What is the earliest and latest time stamp?

> Earliest: 2020-11-03 03:31:30 UTC

> Latest: 2023-01-23 12:29:56 UTC

# Task 2 - Remove Duplicates

1. How many entries are duplicates? Total number and proportion.

> Total = 3,112,100. Proportion $\approx$ 0.0566. 

2. How many rows does your data set have after deleting duplicates?

> 51,848,861

# Task 3 - Descriptive statistics

> The following statistics were calculated using a random sample of 10% of total users in the dataset.

```{r}
#| label: load summary stat dfs
load("data/sample_summary_stats.RData")
```

1. How long are typical journeys? Actual time and actions.

```{r}
#| include: true

# have to remove underscores to render with latex

clean_journey <- journey_stats |>
	rename_with(~ str_replace_all(.x, "_", " ") |> str_to_title())

if (knitr::is_html_output()) {
	gt::gt(clean_journey)
} else {
	tinytable::tt(clean_journey)
}
```

2. What are common actions? Link the action IDs with the actual action
descriptions to answer this question.

```{r}
#| include: true
clean_actions <- common_actions |>
	mutate(event_name = str_replace_all(event_name, "_", " ")) |>
	rename_with(~ str_replace_all(.x, "_", " ") |> str_to_title())

if (knitr::is_html_output()) {
	gt::gt(clean_actions)
} else {
	tinytable::tt(clean_actions)
}
```

3. How much time is between an individual user’s actions?

> *Time is calculated in minutes.*

```{r}
#| include: true
clean_time_btw <- time_between_stats |>
	rename_with(~ str_replace_all(.x, "_", " ") |> str_to_title())

if (knitr::is_html_output()) {
	gt::gt(clean_time_btw)
} else {
	tinytable::tt(clean_time_btw)
}
```
